---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r,echo=TRUE}
library(ggplot2)
activity<-read.csv("activity.csv")
```


## What is mean total number of steps taken per day?


```{r, echo=TRUE}
#steps by day
steps_day<-aggregate(steps~date, data=activity, sum)

#histogram
qplot(steps, data=steps_day,main="Steps by day");

#mean
mean_steps<-mean(steps_day$steps)
mean_steps

#median
median_steps<-median(steps_day$steps)
median_steps
```

The mean steps taken by day are 10766.19 and the median 10765


## What is the average daily activity pattern?

```{r, echo=TRUE}
#Average number of steps taken on the 5 minutes interval
activity_day<-aggregate(steps~interval, data=activity, mean)

#Plot
ggplot(activity_day,aes(interval,steps))+geom_line()+labs(title="Steps by 5 minutes interval")



#Interval the maximum number of steps
max_index<-which.max(activity_day$steps)
max_interval<-activity_day$interval[max_index]
max_interval
```
The maximum number of steps by day are in the 835 interval


## Imputing missing values

```{r,echo=TRUE}

#Number of missing values
missing_values<-sum(!complete.cases(activity))
missing_values

#The missing values are filled with the mean of the steps this interval
new_activity<-activity
for (i in 1:nrow(new_activity)){
  new_activity[i,][is.na(new_activity[i,])]<-activity_day[activity_day$interval==new_activity[i,3],]$steps
}

#Repeat the calculous of the histogram, mean and median
#steps by day
new_steps_day<-aggregate(steps~date, data=new_activity, sum)

#histogram
qplot(steps, data=new_steps_day,main="Steps by day without ");

#mean
new_mean_steps<-mean(new_steps_day$steps)
new_mean_steps

#median
new_median_steps<-median(new_steps_day$steps)
new_median_steps

#difference with before
mean_steps-new_mean_steps
median_steps-new_median_steps

```

The mean of the two datasets (with and without NAs) is the same but the median is 1.19 less in the old dataset.

## Are there differences in activity patterns between weekdays and weekends?

```{r, echo=TRUE}
#New factor variable indicating if it's weekday or weekend
new_activity$date<-as.Date(new_activity$date)
l <- weekdays(new_activity$date) == "sábado" | weekdays(new_activity$date) == "domingo"
new_activity$day[l] <- "weekend"
new_activity$day[!l] <- "weekday"
new_activity$day <- factor(new_activity$day)

#Average number of steps by interval and weekday/weekend
activity_week<-aggregate(steps~interval+day, data=new_activity, mean)

#Plot
ggplot(activity_week,aes(interval,steps))+geom_line()+labs(title="Steps by 5 minutes interval in weekdays/weekend")+facet_grid(day~.)

```

